Spin canonical divisors of log stacky curves

### 2015-06-25 stacky_write_up/spin_cring.tex
add transitions to section 3

\section{Inductive Lemmata}
\label{sec:induction}
First we present the general inductive Lemmata that will be applied
for all genera. We offer various inductive Lemmata that will allow
us to add fractional point $\frac{\alpha}{\beta}P$ to a $\BQ$
 divisor $D$ of $X$ and describe the new generators and relations
of $R_{D+p}$ not in $R_D$. Next, we provide a method specific to
the case of stacky canonical divisors with no log component, which
under certain conditions allows us to transfer information about
the generators and relations of the half canonical ring of a stacky
curve $\sx$' with signature $(g; e_1', \ldots, e_\tau'; 0)$ to a
stacky curve $\sx$ with signature $(g; e_1', \ldots, e_\tau'; 0)$
where each $e_i\in \{e_i', e_i' + 2\}$.

@@

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add_pts}
In this section, we prove various inductive statements under the following setup:
Let $\halfcan'$ be $\mathbb{Q}$-divisor of $X$ (and often a log half canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ not in the support of $\halfcan$.\todo{do we always assume this?}  Then we will explicitly describe elements that minimally generate $R_\halfcan$ over $\halfcan$' and initial terms of relations that minimally generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Grobner Basis).  The proofs of these Lemmas all follow similar patterns, so we will present a few \todo{give the precise number we show} and leave the rest to the reader.

\todo{definition and basic properties of best lower approximations}

\begin{lem}
\label{lem:deg1-sat-ind}
Let $X$ be a genus $g$ curve with a $\mathbb{Q}$-divisor $D$ (as
defined by O'Dorney \todo{reference this-Section 1 of Evan's paper, or preferably define it in the background section})
satisfying $h^0(X, \lfloor{D}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $dD$ for all $d\in \mathbb{N}^+$ meaning that $R_D$ has generically chosen generators (in the appropriate respective degrees)
$u, x_1, \ldots, x_m$ all with $\ord_P(u) = 0$ \todo{Aaron: Instead of having $\deg u = 0$, don't we only need that $\ord_P(u) = \lfloor D|_P \rfloor$?} and $\deg(u)= 1$. Let $D$ be generated by $u, x_1, \ldots, x_m$ in degree at most $\bar{d}$ with relations generated
in degree at most $\bar{\tau}$, where all $x_i$ also satisfy $\ord_P(u)(x_i)$. \todo{Aaron: What does $\ord_P(u)(x_i)$ mean?}  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \text{ for all } d \in \mathbb{
	N} \text{ such that } d \ge \frac{\beta}{\alpha}.
\end{equation}
\label{eqn:deg1-sat-ind}
Let 
\[
	0<\frac{c_1}{d_1}<\ldots<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the non-negative best lower approximation.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1, \ldots, y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

@@

\begin{proof}
Observe that $u_D, x_1, \ldots, x_m$ each of which has no pole at $P$.
These can be chosen generically to not have a zero at $P$ \todo{
check that we can actually do this, or eliminate if it turns out to
be unnecessary}.

Begin by focusing on the $d^{\text{th}}$ component of $R_{D'}$: $(R_
{D'})_d = H^0(X, \lfloor D + \frac{\alpha}{\beta} P \rfloor)$. By
our hypothesis (Equation ~\ref{eqn:deg1_sat_ind_dim}), for any $d
\in \mathbb{N}$ such that $\lfloor d \frac{ \alpha}{\beta} \rfloor > 0$,

@@

>>>>

\section{Inductive Lemmata}
\label{sec:induction}
First we present the general inductive Lemmata that will be applied
for all genera. We offer various inductive Lemmata that will allow
us to add fractional point $\frac{\alpha}{\beta}P$ to a $\BQ$
 divisor $D$ of $X$ and describe the new generators and relations
of $R_{D+p}$ not in $R_D$. Next, we provide a method specific to
the case of stacky canonical divisors, which
under certain conditions allows us to transfer information about
the generators and relations of the half canonical ring of a stacky
curve $\sx$' with signature $(g; e_1', \ldots, e_\tau'; 0)$ to a
stacky curve $\sx$ with signature $(g; e_1, \ldots, e_\tau; 0)$
where each $e_i\in \{e_i', e_i' + 2\}$.

@@

\ssec{Inductive Lemmata: Adding Points}
\label{ssec:add_pts}
In this section, we prove various inductive statements under the following setup:
Let $\halfcan'$ be $\mathbb{Q}$-divisor of $X$ (and often a log half canonical divisor of some $\sx'$), satisfying certain conditions about $\deg(\lfloor d \halfcan'\rfloor)$ for various values of $d\in\mathbb{N}$ and certain saturation conditions. Let $\halfcan=\halfcan'+P$ for some $P$ possibly in the support of $\halfcan$.  Then we will explicitly describe elements that minimally generate\todo{check minimal generation actually works} $R_\halfcan$ over $\halfcan$' and initial terms of relations that minimally generate $\initial(I')$ over $\initial(I)$ (allowing the construction of a corresponding Grobner Basis).  The proofs of these Lemmas all follow similar patterns, so we will present a few \todo{give the precise number we show} and leave the rest to the reader.

First, let's introduce some notation from O'Dorney.   \todo{cite O'downey}
\begin{defn}\label{den:lower-approximation}
If $\frac{\alpha}{\beta}\in \mathbf{Q}$ (for $\alpha\in \mathbf{Z}, \beta\in \mathbf{N}$ written in reduced form), then a \begin{bf}{best lower approximation}\end{bf} to $\frac{\alpha}{\beta}$ is a a fraction $\frac{c}{d}$ (written in reduced form with $d>0$) such that there is no $\frac{c'}{d'}$ with $0<d'<d$ and $\frac{c}{d}\le \frac{c'}{d'}$.  Moreover, the non-negative best lower approximations of $\frac{\alpha}{\beta}$ form a sequence
\[
	0=\frac{c_0}{d_0} < \frac{c_1}{d_1} < \ldots < \frac{c_r}{d_r} = \frac{\alpha}{\beta}.
\]
\end{defn}
\todo{add a picture of lattice}

O'Dorney uses best lower approximations to provide a minimal presentation for the canonical ring of a (not-necessarily effective) divisor $D=\frac{\alpha_1}{\beta_1}P_1+\frac{\alpha_2}{\beta_2}P_2$ of $\mathbf{P}^1$ consisting of at most 2 points in its support, showing that it is generated in degree $e=\max{\beta_1,\beta_2}$ with relations generated in degrees at most $2e$.  He further extends the result to more points, but only bounds the degree of minimal generation by a value on the order the least common multiple of the denominators of fractional points.

The following lemma provides an inductive method to extend O'Dorney's result on $\mathbf{P}^1$ to divisors $D=\sum_i \frac{\alpha_i}{\beta_i}P_i$ consisting of arbitrarily many points satisfying $H^0(X,\lfloor D\rfloor)>0$, showing that they are generated in degrees at most $e=\max(\beta_i)$ with relations generated in degrees at most $2e$.  Furthermore, given some constraints on $H^0(X, \lfloor dD\rfloor)$ for various values of $d$, we can apply the same Riemann-Roch arguments from $\mathbf{P}^1$ in arbitrary genus, allowing a similar inductive result in arbitrary genus $e=\max(\beta_i)$.  To do this, we extend an inductive result by Voight and Zurich-Brown on stacky canonical rings to arbitrary $\mathbf{Q}$-divisors. \todo{cite theorem 8.3.1}

\begin{lem}
\label{lem:deg1-sat-ind}
Let $X$ be a genus $g$ curve with a $\mathbb{Q}$-divisor $\halfcan'$
satisfying $h^0(X, \lfloor{\halfcan'}\rfloor)\ge 1$. Suppose $P$ is not a base-point of $d\halfcan'$ for all $d\in \mathbb{N}^+$ meaning that $R_{\halfcan'}$ has generically chosen generators (in the necessary degrees)
$u, x_1, \ldots, x_m$ with $\ord_P(x_i)=0\; \forall x_i$, $\ord_P(u) = 0$, \todo{Aaron: Instead of having $\deg u = 0$, don't we only need that $\ord_P(u) = \lfloor D|_P \rfloor$?} and $\deg(u)= 1$ when viewed as sections of $\so_D$.  Suppose $u, x_1, \ldots, x_m$ generate $R_D$ in degree at most $\bar{d}$ with relations generated
in degree at most $\bar{\tau}$.  Suppose $D' = D + \frac{\alpha}{\beta} P$
for some $\alpha,\beta \in \mathbb{N}$ such that
\begin{equation}
\label{eqn:deg1_sat_ind_dim}
	h^0(X, \lfloor{dD + d\frac{\alpha}{\beta} P \rfloor}) = h^0(X,\lfloor dD
	\rfloor) + \lfloor d\frac{\alpha} {\beta} \rfloor \text{ for all } d \in \mathbb{
	N} \text{ such that } d \ge \frac{\beta}{\alpha}.
\end{equation}
\label{eqn:deg1-sat-ind}
Let 
\[
	0<\frac{c_1}{d_1}<\ldots<\frac{c_n}{d_n}=\frac{\alpha}{\beta}
\]
be the positive best lower approximations of $\frac{\alpha}{\beta}$.
Then 
\begin{itemize}
\item $R_{D'}$ is generated over $R_D$ by elements $y_1, \ldots, y_n$ such that $y_i$ lies in  
degree $d_i$ and has a pole at $P$ of degree $c_i$

@@

\begin{proof}
Inside of $(R_
{D'})_d = H^0(X, \lfloor D + \frac{\alpha}{\beta} P \rfloor)$. By
our hypothesis (Equation ~\ref{eqn:deg1_sat_ind_dim}), for any $d
\in \mathbb{N}$ such that $\lfloor d \frac{ \alpha}{\beta} \rfloor > 0$,

@@

\begin{rem}\label{rem:sat-1-gen-lem-generic}
The generators in the previous lemma are generic if $\frac{\alpha}{\beta}\le 1$ (so there is not more than one $d_i=1$).  When $\frac{\alpha}{\beta}$, the choice of generators in degrees above 1 is generic, and we can make the choice in degree 1 generic by choosing $\lfloor \frac{\alpha}{\beta}\rfloor$ linearly independent elements in degree 1 with pole at $P$ of order $\lfloor \frac{\alpha}{\beta}\rfloor$ rather than elements with poles of order $1, \ldots, \lfloor \frac{\alpha}{\beta}\rfloor$.
\end{rem}
